Optimizing Abstract Abstract Machines

### 2012-11-13 paper/paper.tex
Compact state graph figure.

\begin{figure}
\hrule
\vspace{2mm}
\begin{alltt}
(define (id x) x)
(define (f n)
  (cond [(<= n 1)  1]
        [else      (* n (f (- n 1)))]))
(define (g n)
  (cond [(<= n 1)  1]
        [else      (+ (* n n) (g (- n 1)))]))
(print (+ ((id f) 3) ((id g) 4)))
\end{alltt}
\hrule
\begin{center}
\begin{tabular}{ccc}
\raisebox{1ex-\height}{
\includegraphics[height=4in]{introspective-base.pdf}}
&
\raisebox{1ex-\height}{
\includegraphics[height=3.7in]{introspective-lazy.pdf}}
&
\raisebox{1ex-\height}{
\includegraphics[height=2.2in]{introspective-lazyc.pdf}}
\\
(a) Baseline
&
(b) Lazy
&
(c) Compiled
\end{tabular}
\end{center}
\caption{Example state graphs for the program above.  Part (a) shows
  the result of the baseline analyzer.  It has long ``corridor''
  transitions and ``diamond'' subgraphs that fan-out from
  nondeterminism and fan-in from joins.  Part (b) shows the result of
  performing nondeterminism lazily and thus avoids many of the diamond
  subgraphs.  Part (c) shows the result of abstract compilation that
  removes intepretive overhead in the form of intermediate states,
  thus minimizing the corridor transitions.  The end result is a more
  compact abstraction of the program that can be generated faster.}
\end{figure}

\newpage
\
\newpage
\section{An abstract$^{\text 2}$ machine for ISWIM}
\label{sec:aam}

@@

Figure~\ref{fig:syntax} defines the (abstract) syntax of ISWIM.
Figure~\ref{fig:aam} defines the semantics of ISWIM as a machine
model.  Evaluation is defined as the set of states reachable by the
reflexive, transitive closure of the machine transition relation.  The
machine is a very slight variation on a standard abstract machine for
ISWIM in ``eval, continue, apply''
form~\cite{dvanhorn:Danvy:DSc}.\marginpar{There's probably a better
  (more specific) cite for eval/apply/continue.}  It can be
systematically derived from a definitional interpreter through a
continuation-passing style transformation and defunctionalization, or
from a structural operational semantics using the refocusing
construction of Danvy and
Nielsen~\cite{dvanhorn:Danvy-Nielsen:RS-04-26}.

>>>>

\begin{figure}
\hrule
\vspace{2mm}
\begin{alltt}
(define (id x) x)
(define (f n)
  (if (<= n 1) 1 (* n (f (- n 1)))))
(define (g n)
  (if (<= n 1) 1 (+ (* n n) (g (- n 1)))))
(print (+ ((id f) 3) ((id g) 4)))
\end{alltt}
\hrule
\begin{center}
\begin{tabular}{ccc}
\raisebox{1ex-\height}{
\includegraphics[height=4in]{introspective-base.pdf}}
&
\raisebox{1ex-\height}{
\includegraphics[height=3.7in]{introspective-lazy.pdf}}
&
\raisebox{1ex-\height}{
\includegraphics[height=2.2in]{introspective-lazyc.pdf}}
\\
(a) Baseline
&
(b) Lazy
&
(c) Compiled (\& lazy)
\end{tabular}
\end{center}
\caption{Example state graphs for the program above.  Part (a) shows
  the result of the baseline analyzer.  It has long ``corridor''
  transitions and ``diamond'' subgraphs that fan-out from
  nondeterminism and fan-in from joins.  Part (b) shows the result of
  performing nondeterminism lazily and thus avoids many of the diamond
  subgraphs.  Part (c) shows the result of abstract compilation that
  removes intepretive overhead in the form of intermediate states,
  thus minimizing the corridor transitions.  The end result is a more
  compact abstraction of the program that can be generated faster.}
\end{figure}

\section{An abstract$^{\text 2}$ machine for ISWIM}
\label{sec:aam}

@@

Figure~\ref{fig:syntax} defines the (abstract) syntax of ISWIM.
Figure~\ref{fig:aam} defines the semantics of ISWIM as a machine
model.  Evaluation is defined as the set of states reachable by the
reflexive, transitive closure of the machine transition relation.  The
machine is a very slight variation on a standard abstract machine for
ISWIM in ``eval, continue, apply''
form~\cite{dvanhorn:Danvy:DSc}.  It can be
systematically derived from a definitional interpreter through a
continuation-passing style transformation and defunctionalization, or
from a structural operational semantics using the refocusing
construction of Danvy and
Nielsen~\cite{dvanhorn:Danvy-Nielsen:RS-04-26}.
